Text embedding models yield high-resolution insights into conceptual knowledge from short multiple-choice quizzes

### 2025-01-04 paper/main.tex
make "proportion-correct" hyphenation consistent throughout

With these ideas in mind, we asked: conditioned on answering a question
correctly, what proportion of all questions (within some radius, $r$, of that
question's embedding coordinate) were answered correctly? We plotted this
proportion as a function of $r$. Similarly, we could ask, conditioned on
answering a question incorrectly, how the proportion of correct responses
changed with $r$. As shown in Figure~\ref{fig:smoothness}, we found that quiz
performance falls off smoothly with distance, and the ``rate'' of the falloff
does not appear to change across the different quizzes, as measured by the
distance at which performance becomes statistically indistinguishable from a
simple proportion correct score (see~\nameref{subsec:smoothness}). This
suggests that, at least within the region of text embedding space covered by
the questions our participants answered (and as characterized using our topic
model), the rate at which knowledge changes with distance is relatively
constant, even as participants' overall level of knowledge varies across
quizzes or regions of the embedding space.

@@

Next, for each participant in the current subsample, and for each of the three
quizzes they completed (separately), we iteratively treated each of the 15
questions appearing on the given quiz as the ``reference'' question. We
constructed a series of concentric 15-dimensional ``spheres'' centered on the
reference question's embedding space coordinate, where each successive sphere's
radius increased by 0.01 (correlation distance) between 0 and 2, inclusive
(i.e., tiling the range of possible correlation distances with 201 spheres in
total). We then computed the proportion of questions enclosed within each
sphere that the participant answered correctly, and averaged these per-radius
proportion correct scores across reference questions that were answered
correctly, and those that were answered incorrectly. This resulted in two
number-of-spheres sequences of proportion-correct scores for each subsample
participant and quiz: one derived from correctly answered reference questions,
and one derived from incorrectly answered reference questions.

We computed the across-subsample-participants mean proportion correct for each
radius value (i.e., sphere) and ``correctness'' of reference question. This
yielded two sequences of proportion-correct scores for each quiz, analogous to
the blue and red lines displayed in Figure~\ref{fig:smoothness}A, but for the
present subsample. For each quiz, we then found the minimum distance from the
reference question (i.e., sphere radius) at which each of these two sequences
of per-radius proportion correct scores intersected the 95\% confidence
interval for the overall proportion correct (i.e., analogous to the black error
bands in Fig.~\ref{fig:smoothness}A).

>>>>

With these ideas in mind, we asked: conditioned on answering a question
correctly, what proportion of all questions (within some radius, $r$, of that
question's embedding coordinate) were answered correctly? We plotted this
proportion as a function of $r$. Similarly, we could ask, conditioned on
answering a question incorrectly, how the proportion of correct responses
changed with $r$. As shown in Figure~\ref{fig:smoothness}, we found that quiz
performance falls off smoothly with distance, and the ``rate'' of the falloff
does not appear to change across the different quizzes, as measured by the
distance at which performance becomes statistically indistinguishable from a
simple proportion-correct score (see~\nameref{subsec:smoothness}). This
suggests that, at least within the region of text embedding space covered by
the questions our participants answered (and as characterized using our topic
model), the rate at which knowledge changes with distance is relatively
constant, even as participants' overall level of knowledge varies across
quizzes or regions of the embedding space.

@@

Next, for each participant in the current subsample, and for each of the three
quizzes they completed (separately), we iteratively treated each of the 15
questions appearing on the given quiz as the ``reference'' question. We
constructed a series of concentric 15-dimensional ``spheres'' centered on the
reference question's embedding space coordinate, where each successive sphere's
radius increased by 0.01 (correlation distance) between 0 and 2, inclusive
(i.e., tiling the range of possible correlation distances with 201 spheres in
total). We then computed the proportion of questions enclosed within each
sphere that the participant answered correctly, and averaged these per-radius
proportion-correct scores across reference questions that were answered
correctly, and those that were answered incorrectly. This resulted in two
number-of-spheres sequences of proportion-correct scores for each subsample
participant and quiz: one derived from correctly answered reference questions,
and one derived from incorrectly answered reference questions.

We computed the across-subsample-participants mean proportion correct for each
radius value (i.e., sphere) and ``correctness'' of reference question. This
yielded two sequences of proportion-correct scores for each quiz, analogous to
the blue and red lines displayed in Figure~\ref{fig:smoothness}A, but for the
present subsample. For each quiz, we then found the minimum distance from the
reference question (i.e., sphere radius) at which each of these two sequences
of per-radius proportion-correct scores intersected the 95\% confidence
interval for the overall proportion correct (i.e., analogous to the black error
bands in Fig.~\ref{fig:smoothness}A).
